---
title: "README"
author: "Tasmyn Mann"
date: "11/02/2020"
output: html_document
---
# Purpose
This document will outline the steps 
I have taken in order to create the fantastic graphs in my 
descriptive paper on changing EU-China trade dynamics and 
their impact on South Africa.

The final paper can be found under "tradedynamics1.pdf"

#Data
The following data was adapted from the World Bank 
national accounts data and the OECD national accounts data file
from the World Development Inidcators
from: https://datacatalog.worldbank.org/public-licenses#cc-by
All data is recorded anually
1.Exports of goods and services (% of GDP)
2.Imports of goods and services (% of GDP)
3. GDP Per Capita in constant 2010 US dollars
4. Net Trade in goods and services (Bop, current US$)
5. Terms of Trade adjustment in constant LCU 
6. Average annual GDP growth rates

The Real Effective Exchange Rate data was adapted from 
Quantec EsayData froom the Main Economic Indicators database
All data is recorded quarterly 
The REER is an index with 2015 as the base year 

#Required packages 
```{r load packages}
if (!require("dplyr")) install.packages("dplyr")
if (!require("devtools")) install.packages("devtools")
devtools::install_github("Stan125/GREA" )
install.packages("ggplot2") 
install.packages("rlang")
install.packages("lubridate")
install.packages("glue")
install.packages("tidyverse")
install.packages("ggthemes")
install.packages("ggsci")
install.packages("rmsfuns")
install.packages("magick")
install.packages("ggrepel")
install.packages("RColorBrewer")
install.packages("ggpubr")
install.packages("readr")
library(dplyr)
library(devtools)
library(ggplot2)
library(rlang)
library(lubridate)
library(rmsfuns)
library(glue)
library(ggthemes)
library(tidyverse)
library(ggthemes)
library(ggsci)
library(magick)
library(ggrepel)
library(RColorBrewer)
library(ggpubr)
library(readr)
```

#Importing the Datasets
```{r data}
TradeData1 <- read.csv(file = "./data/TradeData1.csv") 
nt <- read.csv("./data/nt.csv")  
reer <- read.csv("./data/reer.csv")
tot <- read.csv("./data/tot.csv")
```

#Creating a theme for Graphs and expanding colour palette
```{r colour}
{#Setting Graph Theme
My_Theme = theme(
  axis.title.x = element_text(size = 10),
  axis.text.x = element_text(size = 7),
  axis.text.y = element_text(size = 7),
  axis.title.y = element_text(size = 10))

# Define number of colors
nb.cols <- 10
mycolors <- colorRampPalette(brewer.pal(8, "Paired"))(nb.cols)
}
```

# Graph 1 - South Africas GDP PC

```{r GDPPC}
{#Isolate South Africa GDP PC from Tradedata1 dataset 
SA.GDPPC <- TradeData1 %>% 
  filter(COUNTRY == "SA", SERIES=="GDP PC (constant 2010 US$)", !is.na(VALUE))

SA.GDPPC.G <- ggplot(data=SA.GDPPC) +geom_line(aes(x=DATE, y=VALUE) ,colour="deepskyblue4")+
  labs(x="Time", y="US$ (thousands)", caption="Data Source: World Bank and OECD Data", title ="South Africa GDP Per Capita")+
  scale_x_continuous(breaks = seq(1960, 2020, by = 5)) +
  scale_y_continuous(breaks = seq(4500, 8000, by = 500))+
  My_Theme+
  annotate("rect", fill = "darkgrey", alpha = 0.5, 
           xmin = 2008, xmax = 2020,
           ymin = -Inf, ymax = Inf) 
print(SA.GDPPC.G)
}
```

# Graph 2 - South Africa's Imports and Exports as % of GDP

```{r SA.EX&IM}
{#Filter data to isolate SA exports and imports 
  TradeData1 %>% filter(COUNTRY=="SA") %>% distinct(SERIES)

  target <- c("EX (% of GDP)" , "IM (% of GDP)")
  SA.Trade <- TradeData1 %>% 
    filter(COUNTRY == "SA" , SERIES %in% target, !is.na(VALUE))
  
  #Graph 2 - Create Graph for SA International Trade 
  SA.Trade.G <- ggplot(data=SA.Trade)+ 
    geom_line(aes(x=DATE, y=VALUE, colour=SERIES), size=0.5)+
    theme(plot.title = element_text(hjust=0.5), axis.text.x =element_text(angle=90,hjust=1), legend.position = "bottom")+
    labs(x="Time", y="% of GDP" ,title="South Africa - Exports and Imports as % GDP", 
         caption = "Data Source : World Bank Data")+
    scale_x_continuous(breaks = seq(1960, 2020, by = 5)) +
    scale_y_continuous(breaks = seq(15, 40, by = 5))+
    scale_fill_manual(values = mycolors)+
    My_Theme
  print(SA.Trade.G)}
```

#Graph 3 - South Africas Net Trade 

```{r SAnettrade}
{#Isolate Net Trade into a dataset
  target7 <- c("SA NT")
  ntsa.df <- nt %>% 
    filter(codes %in% target7, !is.na(values))
  #create new column to work with smaller y values on the graph 
  ntsa.df$v <- as.numeric(as.character(ntsa.df$values)) / 100
  #create net trade graph 
  ntsa.G <- ggplot(data= ntsa.df[which(ntsa.df$dates>"1995"),])+ #data only begins in 1995
    geom_line(aes(x=dates, y=v), colour="deepskyblue4")+
    labs(x="Time", y="Billions (US$)" ,title="South Africa - Net Trade in Goods and Services", 
         caption = "Data Source : World Bank and OECD Data")+
    scale_x_continuous(breaks = seq(1996, 2020, by = 2)) +
    scale_y_continuous(breaks = seq(-10000, 5000, by = 2000))+
    scale_fill_manual(values = "deepskyblue4")+
    geom_hline(yintercept=0, colour="darkgrey")
  My_Theme
  ntsa.G <- ntsa.G + guides(fill=guide_legend(title="Country"))#change legend title
  print(ntsa.G)
}
```

#Graph 4 - SA Export composition 2017

```{r SA Export Comp 2017}
  {
    #create the data frame from OEC Data 
  ExportComp <- data.frame(
  Catergory = c("Precious Metals", "Mineral Products", "Metals",
                "Transportation", "Machines", "Chemical Products",
                "Vegetable Products", "Foodstuffs","Animal Products","Other"),
  Composition = c(34, 19, 11, 10, 5.6, 4.9, 4.7, 2.6, 1.5, 6.7)
)
head(ExportComp)

#Labels and cumsum for pie chart
ExportComp <- ExportComp %>%
  arrange(desc(Catergory)) %>%
  mutate(lab.ypos = cumsum(Composition) - 0.5*Composition)
ExportComp

  #creating Pie Chart on ggplot
ggplot(data=ExportComp, 
       aes(x="", y= Composition ,fill = Catergory))+
  geom_bar(width = 1, stat="identity")+
  coord_polar("y" , start=0) +
  geom_text(x=1.6, aes(y=lab.ypos, label = paste0(Composition, "%")),
            col="black") +
  scale_fill_manual(values = mycolors)+
  theme_void()+
  labs(fill="Catergory", x=NULL, y=NULL, title="South Africa's Export Goods Composition (2017)",
       axis.text.x = element_text(size = 2))
}
```

#Graph 5 - Export Composition 1995

```{r SA Export Comp 1995}
 {
#create 1995 data frame 
   ExportComp1995 <- data.frame(
  Catergory1 = c("Precious Metals", "Mineral Products", "Metals",
                 "Transportation", "Machines", "Chemical Products",
                 "Vegetable Products", "Foodstuffs","Animal Products","Other"),
  Composition1 = c(12, 23, 20, 2.9, 6, 5.8, 6.8, 5.3, 3.1, 15.1)
)
head(ExportComp1995)

ExportComp1995 <- ExportComp1995 %>%
  arrange(desc(Catergory1)) %>%
  mutate(lab.ypos = cumsum(Composition1) - 0.5*Composition1)
ExportComp1995

ggplot(data=ExportComp1995, 
       aes(x="", y= Composition1 ,fill = Catergory1))+
  coord_polar("y" , start=0) +
  geom_text(x=1.6, aes(y=lab.ypos, label = paste0(Composition1, "%")),
            col="black") +
  geom_bar(width = 1, stat="identity")+
  scale_fill_manual(values = mycolors)+
  theme_void()+
  labs(fill="Catergory1", x=NULL, y=NULL, title="South Africa's Export Goods Composition (1995)",
       axis.text.x = element_text(size = 2))
}
```

Graph 6-  South Africa Terms of Trade 

```{r SA TOT}
 {
#replacing N/A values with 0
tot[is.na(tot)] = 0
#Graph
tot.G <- ggplot(data=tot) +geom_line(aes(x=YEAR, y=TOT) ,colour="deepskyblue4")+
  labs(x="Time", y="LCU (Billions)", caption="Data Source: World Bank and OECD Data", title ="South Africa - Terms of Trade")+
  scale_x_continuous(breaks = seq(1960, 2020, by = 10)) +
  scale_y_continuous(breaks = seq(-200, 200, by = 50))+
  scale_fill_manual(values = mycolors)+
  geom_hline(yintercept=0, colour="darkgrey")+
  My_Theme
print(tot.G)
}
```

#Graph 7 - GDP Growth for all 3 regions

```{r GDP Growth}
 {
#GDP Growth Graph 
target5 <- c("SA GDP G", "CHN GDP G", "EU GDP G")
gdp.df <- nt %>% 
  filter(codes %in% target5, !is.na(values))

gdpg.G <- ggplot(data= gdp.df[which(gdp.df$dates>"1970"),])+ 
  geom_line(aes(x=dates, y=values, colour=codes), size=0.5)+
  theme(plot.title = element_text(hjust=0.5), 
        axis.text.x = element_text(angle=90,hjust=1),
        legend.position = "bottom")+
  labs(x="Time", y="GDP Growth (%)" ,title="Average Annual Growth", 
       caption = "Data Source : World Bank and OECD Data")+
  scale_x_continuous(breaks = seq(1960, 2020, by = 5)) +
  scale_fill_manual(values = mycolors)+
  My_Theme
gdpg.G <- gdpg.G + guides(fill=guide_legend(title="Country"))
print(gdpg.G) 
}
```

Graph 8 - Chinas Exports and Imports 

```{r China Ex and Im}
 {
#Isolate Imports and Exports for China
target1 <- c("EX (% of GDP)" , "IM (% of GDP)")
CHN.Trade <- TradeData1 %>% 
  filter(COUNTRY == "CHN" , SERIES %in% target1, !is.na(VALUE))

#Create Graph for China International Trade 
CHN.Trade.G <- ggplot(data=CHN.Trade)+ 
  geom_line(aes(x=DATE, y=VALUE, colour=SERIES), size=0.5)+
  theme(plot.title = element_text(hjust=0.5), axis.text.x =element_text(angle=90,hjust=1), legend.position = "bottom")+
  labs(x="Time", y="% of GDP" ,title="China - Exports and Imports as % GDP", 
       caption = "Data Source : World Bank")+
  scale_x_continuous(breaks = seq(1960, 2020, by = 5)) +
  scale_y_continuous(breaks = seq(0, 40, by = 5))+
  scale_fill_manual(values = mycolors)+
  My_Theme
print(CHN.Trade.G)
}
```

#Graph 9 - EU Imports and Exports
```{r EU Ex and Im}
 {
#Filter Data to isolate Europe exports and imports
target2 <- c("EX (% of GDP)" , "IM (% of GDP)")
EU.Trade <- TradeData1 %>% 
  filter(COUNTRY == "EU" , SERIES %in% target2, !is.na(VALUE))

#Create Graph for EU International Trade 
EU.Trade.G <- ggplot(data=EU.Trade)+ 
  geom_line(aes(x=DATE, y=VALUE, colour=SERIES), size=0.5)+
  theme(plot.title = element_text(hjust=0.5), axis.text.x =element_text(angle=90,hjust=1), 
        legend.position = "bottom")+
  labs(x="Time", y="% of GDP" ,title="EU - Exports and Imports as % GDP", 
       caption = "Data Source : World Bank")+
  scale_x_continuous(breaks = seq(1960, 2020, by = 5)) +
  scale_y_continuous(breaks = seq(0, 50, by = 5))+
  scale_fill_manual(values = mycolors)+
  My_Theme
print(EU.Trade.G)
}
```

#Graph 10 - Net Trade of all 3 regions 

```{r Net Trade EU SA and China}
 {
#Create Graph Net Trade 
target4 <- c("EU NT", "CHN NT", "SA NT")
nt.df <- nt %>% 
  filter(codes %in% target4, !is.na(values))

nt.df$v <- as.numeric(as.character(nt.df$values)) / 100

nt.G <- ggplot(data= nt.df[which(nt.df$dates>"1995"),])+ 
  geom_line(aes(x=dates, y=v, colour=codes), size=0.5)+
  theme(plot.title = element_text(hjust=0.5), 
        axis.text.x = element_text(angle=90,hjust=1),
        legend.position = "bottom")+
  labs(x="Time", y="Net Trade (millions US$)" ,title="Net Trade - Goods and Services", 
       caption = "Data Source : World Bank and OECD Data")+
  scale_x_continuous(breaks = seq(1995, 2020, by = 5)) +
  scale_fill_manual(values = mycolors)
nt.G <- nt.G + guides(fill=guide_legend(title="Country"))
print(nt.G)
}
```

#Graph 11- Real Effective Exchange Rates 

```{r REER}
 {
#change the date column as its quarterly data
   reer$date <- as.Date(as.character(reer$date))
target8 <- c("EU REER", "CHN REER", "SA REER")
reer.df <- reer %>% 
  filter(code %in% target8, !is.na(rate))

reer.G <- ggplot(data = reer.df)+ 
  geom_line(aes(x=date, y=rate, colour=code), size=0.5)+
  theme(plot.title = element_text(hjust=0.5), 
        axis.text.x = element_text(angle=90,hjust=1),
        legend.position = "bottom")+
  labs(x="Time", y="REER (%)" ,title="Real Effective Exchange Rate", 
       caption = "Data Source : Quantec")+
  scale_y_continuous(breaks = seq(60, 160, by = 20)) +
  scale_fill_manual(values = mycolors)+
  My_Theme
print(reer.G)
}
```

Graph 12 - Alternative REER Graph 
```{r REER2}
 {
#Graph REER 
reer.sa <- reer %>% 
  filter(code == "SA REER")
reer.sa$date <- as.Date(as.character(reer.sa$date))
reer.sa.G <- ggplot(data=reer.sa) +geom_line(aes(x=date, y=rate, group=1) ,colour="deepskyblue4")+
  labs(x="Time", y="REER(%)")
print(reer.sa.G)

reer.chn <- reer %>% 
  filter(code == "CHN REER")
reer.chn$date <- as.Date(as.character(reer.chn$date))
reer.chn.G <- ggplot(data=reer.chn) +geom_line(aes(x=date, y=rate, group=1) ,colour="deepskyblue4")+
  labs(x="Time", y="REER (%)")
print(reer.chn.G)

reer.eu <- reer %>% 
  filter(code == "EU REER")
reer.eu$date <- as.Date(as.character(reer.eu$date))
reer.eu.G <- ggplot(data=reer.eu) +geom_line(aes(x=date, y=rate, group=1) ,colour="deepskyblue4")+
  labs(x="Time", y="REER (%)")
print(reer.eu.G)

ggarrange(reer.sa.G, reer.chn.G, reer.eu.G, ncol = 3, labels = c("SA", "CHINA", "EU")) 
}
```

